Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-04 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
EvoDrawings,EvoDraw, organization

Data available from the interaction is also used to increase the engegment of 
users in the system by applying  gamification techniques. Gamification is defined by 
Deterding et al. \cite{deterding2011game} as
\begin{quote}
  the use of game design elements in non-game contexts.
\end{quote}  
The gamification element employed in this case study has been a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. Levels depend
on the score and certain features of the game are only available to gamers when 
they reach a giving level. % Say what we are doing in this case - JJ
%%%
%%% Change if case studies change
%In this work, a case study consisting in a web application for evolving artistic drawings
%is presented, providing both conceptual and implementation details of the framework. 
%In order to test the framework three versions of the application are compared to
%measure the degree of relationship between users and their participation,
%when appling a gamification technique. 
In this work, three case studies are presented. First, the development of a web based C-IEC application 
for evolving artistic drawings is described, to give the reader details about 
the utilization of the proposed framework. Then, a rewarding mechanism is 
implemented in the same application, to study the changes in participation,
when appling a gamification technique. Later two other case studies are briefly presented
to show how the framework is applied in other domains.

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of Interactive Evolution. Then, Section \ref{sec:framework} presents the human centered framework of
for C-IEC appliactions which the main proposal of this work. Next three case studies are presented in
sections

An implementation using the ES-I framework 
is detailed in Section \ref{sec:implementation}.
The case study and results are presented in Sections \ref{sec:experiments} and  
\ref{sec:results}. Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

@@

\item {\bf Interfaces}
The appropiate human interface will depend on the application domain, 
the current framework implemantation employs a web based application. Developers 
can use different templates, to present one or more phenotypes 
at a time, and the type of rating system: like or one to five stars. 
Another implementations do not need a graphical interface at all, for example the
EvoDraw instalation or the XYZ Project \cite{de2014artists}.

@@

\section{Case Study: EvoDraw}
As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

@@

\subsection{Fitness Assignment}
\label{sec:assignment}

The ES-I platform has been programmed to employ a 
collaborative methodology for performing fitness assignment. 
Several registered users assign a quality assessment to a single
phenotype and then an aggregated fitness value is calculated,
depending on the rating (from 1 to 5 stars) by each particular user,  
resulting in a many-to-many relationship between users and
phenotypes. Other components or users of the systems can 
query this user-phenotype relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular phenotype, or the user with more participation 
\cite{picbreeder}.
In order to do this, metadata about phenotypes 
must be permanently stored, even
if they are no longer in the active population.

@@

\subsection{Graph Model} 
The graph model for the EvoDraw applications has the following types of
nodes: {\bf User}, {\bf Phenotype} and {\bf Collection} . The Collection node represents
a collection of drawings belonging to users. One collection can contain many phenotypes 
or be empty. A single phenotype could be shared by many collections. The interaction 
between these entities are represented by the following edges:

@@

\subsection{Gamification}
The rewarding mechanism as it is applied in EvoDraw gives more importance 
to the preference of those users with higher reputation
as given by their score points and experience levels.  
Each time a user does on of these actions their score is incremented by one:
\begin{itemize}
\item Start a session.
\item Rate a phenotype.
\item Create a collection.
\item Save a phenotype of the wall to a collection.
\item Save a phenotypes from a friend's collection.
\item Explore collections of other friends.
\end{itemize}

@@

\subsection{Experimental Setup}
\label{sec:experiments}

Three versions of EvoDraw were compared:
\begin{itemize}
\item Base (B). All users have the same weight.
\item Non Graph Gamification (G). Only experience is considered
\item Graph Gamification (GG) . Both experience and participation are considered.
\end{itemize}
When gamification was employed, all score values known where presented to users
and a ranking of users by weight was shown in a window.
The three versions of the EvoDrawing web application were deployed to the
Heroku cloud service with the same configuration: Heroku Free option with a 512 MB of RAM, 
adding only the GrapheneDB and RedisTOGO plug-ins.

Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.

@@

\section{Case Study: EvoDraw-Kinect}

A variation of the previous application is The EvoDraw installation , using the same
framework but with a different form of interaction. Humans interact with the animations 
by simply looking at them. Again the population is stored in the EvoSpace container and
client computers remove animations from the EvoSpace to be presented in a display one at a time. 
After a certain amount of time,  animations are put back to the server along with information 
generated from the interaction. To continue, another is again removed and presented to viewers. 
An important feature is that once an animation is removed from the server, no other client can remove it, 
until the client that remove it puts the animation back, this means that any
animation is been shown only on one client at a given time. This cycle continues until there
is enough interactions to create a new generation of animations.

@@

\section{Case Study: XYZ-Unplugged}

>>>>

Data available from the interaction is also used to increase the engegment of 
users in the system by applying  gamification techniques. Gamification is defined by 
Deterding et al. \cite{deterding2011game} as
\begin{quote}
  the use of game design elements in non-game contexts.
\end{quote}  
The gamification element employed in this case study has been a rewarding mechanism  
\cite{dubois2013understanding}. In general rewards  consist of a reputation system 
with score points, levels and leader boards. Points are awarded to users in response of 
the accomplishment of certain activities that need to be encouraged. Levels depend
on the score and certain features of the game are only available to gamers when 
they reach a giving level. % Say what we are doing in this case - JJ
%%%
%%% Change if case studies change
%In this work, a case study consisting in a web application for evolving artistic drawings
%is presented, providing both conceptual and implementation details of the framework. 
%In order to test the framework three versions of the application are compared to
%measure the degree of relationship between users and their participation,
%when appling a gamification technique. 
In this work, three case studies are presented. First, the development of a web based C-IEC application 
for evolving artistic drawings is described to give the reader details about 
the utilization of the proposed framework. Then, a rewarding mechanism is 
implemented in the same application to study the changes in participation
when applying a gamification technique. Finally, two other case studies are briefly presented
to show how the framework is applied to other domains.

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of interactive evolution. Then, Section \ref{sec:framework} presents the human centered framework of
for C-IEC appliactions which the main proposal of this work. The implementation of the data model
is detailed in Section \ref{sec:implementation}. Next three case studies are presented: EvoDrawings in
Section \ref{sec:evodraw}, EvoDraw Installation in Section\ref{sec:evodraw}, 
and XYZ in Section\ref{sec:evodraw}. Finally, a concluding remarks are provided 
in Section \ref{sec:conclusions}.

@@

\item {\bf Interfaces}
The appropiate human interface will depend on the application domain, 
the current framework implemantation employs a web based application. Developers 
can use different templates, to present one or more phenotypes 
at a time, and the type of rating system: like or one to five stars. 
Another implementations do not need a graphical interface at all, for example the
EvoDrawings instalation or the XYZ Project \cite{de2014artists}.

@@

\section{Case Study: EvoDrawings}
\label{sec:evodraw}
As a case study, a IEA was developed using the 
EvoSpace-Interactive (ES-I) platform \cite{garcia2013evospace}
A brief description of the application is presented next, focusing
on the data elements that where ported to the graph model.

@@

\subsection{Fitness Assignment}
The ES-I platform has been programmed to employ a 
collaborative methodology for performing fitness assignment. 
Several registered users assign a quality assessment to a single
phenotype and then an aggregated fitness value is calculated,
depending on the rating (from 1 to 5 stars) by each particular user,  
resulting in a many-to-many relationship between users and
phenotypes. Other components or users of the systems can 
query this user-phenotype relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular phenotype, or the user with more participation 
\cite{picbreeder}.
In order to do this, metadata about phenotypes 
must be permanently stored, even
if they are no longer in the active population.

@@

\subsection{Graph Model} 
The graph model for the EvoDrawings application has the following types of
nodes: {\bf User}, {\bf Phenotype} and {\bf Collection} . The Collection node represents
a collection of drawings belonging to users. One collection can contain many phenotypes 
or be empty. A single phenotype could be shared by many collections. The interaction 
between these entities are represented by the following edges:

@@

\subsection{Gamification}
The rewarding mechanism as it is applied in EvoDrawings gives more importance 
to the preference of those users with higher reputation
as given by their score points and experience levels.  
Each time a user does on of these actions their score is incremented by one:
\begin{itemize}
\item Start a session.
\item Rate a phenotype.
\item Create a collection.
\item Save a phenotype of the wall to a collection.
\item Save a phenotypes from a friend's collection.
\item Explore collections of other friends.
\end{itemize}

@@

\subsection{Experimental Setup}

Three versions of EvoDrawings were compared:
\begin{itemize}
\item Base (B). All users have the same weight.
\item Non Graph Gamification (G). Only experience is considered
\item Graph Gamification (GG) . Both experience and participation are considered.
\end{itemize}
When gamification was employed, all score values known where presented to users
and a ranking of users by weight was shown in a window.
Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.

@@

\section{Case Study: EvoDraw-Kinect}
\label{sec:kinect}
A variation of the previous application is The EvoDraw installation , using the same
framework but with a different form of interaction. Humans interact with the animations 
by simply looking at them. Again the population is stored in the EvoSpace container and
client computers remove animations from the EvoSpace to be presented in a display one at a time. 
After a certain amount of time,  animations are put back to the server along with information 
generated from the interaction. To continue, another is again removed and presented to viewers. 
An important feature is that once an animation is removed from the server, no other client can remove it, 
until the client that remove it puts the animation back, this means that any
animation is been shown only on one client at a given time. This cycle continues until there
is enough interactions to create a new generation of animations.

@@

\section{Case Study: XYZ-Unplugged}
\label{sec:xyz}
